---
title: "PA1_template.Rmd"
output: html_document
---
# Reproducible Research: Peer Assessment 1
-----
## Code for reading in the dataset and/or processing the data

```{r}
setwd("~/ML/Reproducible Research/WK1/repdata_data_activity")
steps <-read.csv("activity.csv")
stepperday<-tapply(steps$steps, steps$date, sum)
```

##Histogram of the total number of steps taken each day
```{r}
hist(stepperday, main="Steps per day", xla="steps", yla="number of days")
```

##Mean and median number of steps taken each day 
```{r}
mean(stepperday, na.rm=T)
median(stepperday, na.rm=T)
```


##Time series plot of the average number of steps taken
```{r}
stepmin<-aggregate(steps$steps, list(steps$interval), mean, na.rm=T)
names(stepmin)<-c("interval", "steps")
plot(stepmin$interval, stepmin$steps, type="l")
```

##The 5-minute interval that, on average, contains the maximum number of steps
```{r}
stepmin[order(stepmin$steps),][288,]$interval
```

## Code to describe and show a strategy for imputing missing data

Some timepoints are zero, so we need to determine whether a give point should be zero, and if not, impute based on interval and day. 

```{r}
length(which(is.na(steps$steps)))
steps$hour = floor(steps$interval/100)
steps$period=floor(steps$hour/3)
steps$period<-factor(steps$period)
levels(steps$period)<-c("0-2", "3-5", "6-8", "9-11", "12-14", "15-17", "18-20", "21-23")
mod<-lm(steps ~ period, data=steps)
mod
steps$stepsi<-steps$steps
steps$stepsi[is.na(steps$steps)]<-predict(mod, newdata=steps[is.na(steps$steps),])
```

## Histogram of the total number of steps taken each day after missing values are imputed
```{r}
stepperdayi<-tapply(steps$stepsi, steps$date, sum, na.rm=T)
stepperdayi
hist(stepperdayi, main="Steps per day (with imputed data)", xla="steps", yla="number of days", col="#ff99ff")
```
*mean (with imputed data)*:
```{r}
mean(stepperdayi, na.rm=T)
```
*median  (with imputed data)*:
```{r}
median(stepperdayi, na.rm=T)
```



## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
steps$ddate<-as.character(steps$date)
steps$ddate<-as.Date(steps$ddate, format="%Y-%m-%d")
steps$weekday<-weekdays(steps$ddate)
steps$weekend<-F
steps$weekend[steps$weekday %in% c("Saturday", "Sunday")]<-T

stepmin.i.weekdays<-aggregate(steps$stepsi[!steps$weekend], list(steps$interval[!steps$weekend]), mean, na.rm=T)
stepmin.i.weekends<-aggregate(steps$stepsi[steps$weekend], list(steps$interval[steps$weekend]), mean, na.rm=T)
names(stepmin.i.weekdays)<-c("interval", "steps")
names(stepmin.i.weekends)<-c("interval", "steps")


par(mfrow = c(2,1), mar = c(4, 4, 2, 1), oma = c(0, 0, 2, 0))
plot(stepmin.i.weekends$interval, stepmin.i.weekends$steps, pch="", ylab="Steps", xlab="", main="weekend", type="l", ylim=c(0,220), col="blue")
plot(stepmin.i.weekdays$interval, stepmin.i.weekdays$steps, pch="", ylab="Steps", xlab="", main="weekday", type="l",  ylim=c(0,220), col="darkred")
```